---
title: "LinearRegression_MedicalExp"
author: "Ashley Bischof"
---

Use linear regression to predict medial expenses based on demographic data

##Read in and Explore the Data
```{r}
library(ggplot2)

insurance = read.csv("~/Dropbox/R ref/Machine Learning with R/chapter 6/insurance.csv", stringsAsFactors = TRUE)

str(insurance)
#Data inlcudes: age, sex, bmi, children, smoker, region, charges (dependent variable)

summary(insurance$charges)
ggplot(insurance, aes(charges)) + geom_histogram() + theme_minimal()
#the charges data is extremely right skewed 
ggplot(insurance, aes(charges)) + geom_histogram() + theme_minimal() + scale_x_log10()
#logging the data improves this but it is still not very guassian


#determine the equality of the levels in factor variables:
table(insurance$sex) #equal
table(insurance$region) #equal
table(insurance$smoker) #~4x as many nonsmokers

#understand distribution of continuous variables:
ggplot(insurance, aes(age)) + geom_histogram() + theme_minimal()
#fairly well disributed
ggplot(insurance, aes(bmi)) + geom_histogram() + theme_minimal()
#pretty guassian

ggplot(insurance, aes(children)) + geom_histogram() + theme_minimal()
#count variable, extremely right skewed, and strictly non-negative with the highest bar at 0


#take a quick look at correlations between variables
cor(insurance[c("age", "bmi", "children", "charges")])
#no strong correlations but a few week:
#age and bmi; age and charges; bmi and charges, children and charges

#Visualize with a scatterplot matrix
pairs(insurance[c("age", "bmi", "children", "charges")])

library(psych)
pairs.panels(insurance[c("age", "bmi", "children", "charges")])
ggpairs(insurance, columns = c("age", "bmi", "children", "charges")) 

ggplot(insurance, aes(x = sex, y = charges)) + geom_boxplot() + theme_minimal()
ggplot(insurance, aes(x = region, y = charges)) + geom_boxplot() + theme_minimal()
ggplot(insurance, aes(x = factor(children), y = charges)) + geom_boxplot() + theme_minimal()
ggplot(insurance, aes(x = smoker, y = charges)) + geom_boxplot() + theme_minimal()
```

##Fit a Linear Model
```{r}
ins_model = lm(charges ~ ., data = insurance)
ins_model
```

##Evaluate Model Performance
```{r}
summary(ins_model)

#plot residuals
res = resid(ins_model)
fitted = predict(ins_model)
plot(y = res, x = fitted)
plot(y = res, x = insurance$bmi)
hist(res)
```


##Fit the model with non-linear, binary indicators, and interactions included
```{r}
#age squared
insurance$age2 = insurance$age^2

#bmi binary up to overweight (30)
insurance$bmi30 = ifelse(insurance$bmi >= 30, 1, 0)

#smoking and overweight together

ins_model2 = lm(charges ~ age + age2 + children + bmi + sex + bmi30*smoker + region, data = insurance)

summary(ins_model2)


#plot residuals
res2 = resid(ins_model2)
fitted2 = predict(ins_model2)
plot(y = res2, x = fitted2)
plot(y = res2, x = insurance$bmi)
hist(res2)
```


##need to explore logging the data and analyzing the residuals


##Fit the model to log(charges) with non-linear, binary indicators, and interactions included
```{r}
#age squared
insurance$age2 = insurance$age^2

#bmi binary up to overweight (30)
insurance$bmi30 = ifelse(insurance$bmi >= 30, 1, 0)

#smoking and overweight together

ins_model3 = lm(log(charges) ~ age + age2 + children + bmi + sex + smoker*bmi30 + region, data = insurance)

summary(ins_model3)


#plot residuals
res3 = resid(ins_model3)
fitted3 = predict(ins_model3)
plot(y = res3, x = fitted3)
plot(y = res3, x = insurance$bmi)
plot(y = res3, x = insurance$age)
plot(y = res3, x = insurance$children)
plot(y = res3, x = insurance$sex)
plot(y = res3, x = insurance$smoker)
hist(res3)
```


##Fit the model to log(charges) and children as factor with non-linear, binary indicators, and interactions included
```{r}
#age squared
insurance$age2 = insurance$age^2

#bmi binary up to overweight (30)
insurance$bmi30 = ifelse(insurance$bmi >= 30, 1, 0)

#smoking and overweight together

ins_model4 = lm(log(charges) ~ age + age2 + factor(children) + bmi + sex + smoker*bmi30 + region, data = insurance)

summary(ins_model3)


#plot residuals
res4 = resid(ins_model4)
fitted4 = predict(ins_model4)
plot(y = res4, x = fitted4)
plot(y = res4, x = insurance$bmi)
plot(y = res4, x = insurance$age)
plot(y = res4, x = insurance$children)
plot(y = res4, x = insurance$sex)
plot(y = res4, x = insurance$smoker)
hist(res4)
```
