---
title: "NaiveBayes_spam"
author: "Ashley Bischof"
output: html_document
---

#Use Naive Bayes to classify SMS text messages into spam or not:

## Prepare and Clean Data

Pull in Data and have a quick look:
```{r}
sms_raw = read.csv("~/Dropbox/R ref/Machine Learning with R/chapter 4/sms_spam.csv", stringsAsFactors = FALSE)

str(sms_raw)

sms_raw$type = factor(sms_raw$type)
str(sms_raw)

table(sms_raw$type)
```
'data.frame':  5559 obs. of  2 variables:
 $ type: Factor w/ 2 levels "ham","spam": 1 1 1 2 2 1 1 1 2 1 ...
 $ text: chr  "Hope you are having a good week. Just checking in" "K..give back my thanks." "Am also doing in cbe only. But have to pay." "complimentary 4 STAR Ibiza Holiday or £10,000 cash needs your URGENT collection. 09066364349 NOW from Landline not to lose out!"| __truncated__ ...

 ham spam 
4812  747 

Process the texts into single words and remove punctuation and uninteresting words
```{r}
library(tm)

sms_corpus = VCorpus(VectorSource(sms_raw$text))
#build a corpus of documents
#sms_corpus = Corpus(VectorSource(sms_raw$text))

print(sms_corpus)
inspect(sms_corpus[1:3])

#learn more about text importing options
#print(vignette("tm")) 

#CLEAN TEXT DATA
#convert all messages to lowercase
corpus_clean = tm_map(sms_corpus, content_transformer(tolower))
#remove numbers
corpus_clean = tm_map(corpus_clean, content_transformer(removeNumbers))

#remove stop words defined by tm
#stopwords()
corpus_clean = tm_map(corpus_clean, removeWords, stopwords())
#remove punctuation
corpus_clean = tm_map(corpus_clean, content_transformer(removePunctuation))

#remove extra whitespace
corpus_clean = tm_map(corpus_clean, stripWhitespace)

inspect(corpus_clean[1:3])

#tokenization: split the messages into individual components
#create sparse matrix: rows are documents and columns are terms/words
#each cell stores a count of the times the word in the column appears in the doc indicated by the row
sms_dtm = DocumentTermMatrix(corpus_clean)
```



Split the data into a training and test dataset
  random order so take first 75% (4169) for training and last 1390 for test
```{r}
#split raw data
sms_raw_train = sms_raw[1:4169, ]
sms_raw_test = sms_raw[4170:5559,]

#split document-term matrix
sms_dtm_train = sms_dtm[1:4169, ]
sms_dtm_test = sms_dtm[4170:5559, ]

#split corpus
sms_corpus_train = corpus_clean[1:4169]
sms_corpus_test = corpus_clean[4170:5559]


#compare training and test datasets
prop.table(table(sms_raw_train$type))
prop.table(table(sms_raw_test$type))
```

Visualize using word clouds
```{r}
library(wordcloud)

wordcloud(corpus_clean, min.freq = 40, random.order = FALSE)
wordcloud(sms_corpus_train, min.freq = 40, random.order = FALSE)

spam = subset(sms_raw_train, type == 'spam')
ham = subset(sms_raw_train, type == 'ham')

wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 40, scale = c(3, 0.5))
```



Transform the sparse matrix into a data structure to train a naive Bayes classifier
Eliminate words which appear in less than 0.1% of records (5 here)
```{r}
#pick out the words that appear at least 5 times
sms_dict = findFreqTerms(sms_dtm_train, 5)

#pull the words that appear at least 5 times
sms_train = DocumentTermMatrix(sms_corpus_train, list(dictionary = sms_dict))
sms_test = DocumentTermMatrix(sms_corpus_test, list(dictionary = sms_dict))

#convert counts to a factor variable (yes or no) for word appearance
convert_counts = function(x) {
  x = ifelse(x > 0, 1, 0)
  x = factor(x, levels = c(0,1), labels = c("No", "Yes"))
  return(x)
  }

sms_train = apply(sms_train, MARGIN = 2, convert_counts)
sms_test = apply(sms_test, MARGIN = 2, convert_counts)
```



## Training a model on the data
```{r}
library(e1071)
#also check out k1aR NaiveBayes()

#build model on sms_train matrix with laplace = 0
sms_classifier = naiveBayes(sms_train, sms_raw_train$type)
```


##Perform predition and evaluate model performance
```{r}
#predict sms_test group with sms_classifier
sms_test_pred = predict(sms_classifier, sms_test)

#evaluate model performance
library(gmodels)
CrossTable(sms_test_pred, sms_raw_test$type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```


Test model performance with laplace = 1
```{r}
#build model on sms_train matrix with laplace = 1
sms_classifier2 = naiveBayes(sms_train, sms_raw_train$type, laplace = 1)

#predict sms_test group with sms_classifier
sms_test_pred2 = predict(sms_classifier2, sms_test)

#evaluate model performance
CrossTable(sms_test_pred2, sms_raw_test$type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

