---
title: 'Nearest Neighbors: Tumor Histology'
output: html_document
---

Use morphometric analysis of breast cancer H&Es matched with pathologist diagnosis to cluster data using nearest neighbors:


Load data and have a look clean up a bit
```{r}
wbcd = read.csv("~/Dropbox/R ref/Machine Learning with R/chapter 3/wisc_bc_data.csv", stringsAsFactors = FALSE)

str(wbcd)
#remove first column containing ID
wbcd = wbcd[-1]

table(wbcd$diagnosis)
#357 benign (b) and 212 malignant (M) currently being expressed as char
#make diagnosis into factor variable with more descriptive titles
wbcd$diagnosis = factor(wbcd$diagnosis, levels = c("B", "M"), labels = c("Benign", "Malignant"))


table(wbcd$diagnosis)
round(prop.table(table(wbcd$diagnosis)) *100, digits = 1)
```

Take a look at numerical data and clean:
```{r}
summary(wbcd[c("radius_mean", "area_mean", "smoothness_mean")])

# values are very different need to normalize

Std_normalize = function(x)
  {return ((x-min(x)) / (max(x) - min(x)))}

#test normalization functions
test1 = c(1, 2, 3, 4, 5)
test2 = c(10, 20, 30, 40, 50)

Std_normalize(test1)
Std_normalize(test2)

#apply Normalization to all numerical data
wbcd_sn = as.data.frame(lapply(wbcd[2:31], Std_normalize))

summary(wbcd_sn[c("radius_mean", "area_mean", "smoothness_mean")])
```


split data into a training and test (100) dataset
```{r}
#data is already randomized in this case so we can just take the final 100 as our test set
wbcd_train = wbcd_sn[1:469, ]
wbcd_test = wbcd_sn[470:569, ]

wbcd_train_labels = wbcd[1:469, 1]
wbcd_test_labels = wbcd[470:569, 1]
```


First try to train the model and classify with standard min-max normalization 
```{r}
library(class)

#choose k = 21 as this is ~sqrt(469) and odd to avoid ties
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)
```

Evaluate the model performance
```{r}
library(gmodels)

CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
```
CrossTable:



Indicates that there is 2% error, 2 cases were classified as benign when they had clinically been called malignant. 


Try Z normalization
```{r}
#use scale to z normalize the data
wbcd_zn = as.data.frame(scale(wbcd[-1]))

summary(wbcd_zn$area_mean)

#data is already randomized in this case so we can just take the final 100 as our test set
wbcd_train = wbcd_zn[1:469, ]
wbcd_test = wbcd_zn[470:569, ]

#classify
wbcd_test_pred = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

#compare
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)
```
CrossTable:


Z normalization led to a slightly worse result oveall, with only 95% correctly classified



Try testing alternative values of k
```{r}
#assign training and test data from standard normalziation again
wbcd_train = wbcd_sn[1:469, ]
wbcd_test = wbcd_sn[470:569, ]

#classify choose k = 1
wbcd_test_pred_1 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

# classify choose k = 5
wbcd_test_pred_5 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

#choose k = 11
wbcd_test_pred_11 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

#choose k = 15
wbcd_test_pred_15 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

#choose k = 21
wbcd_test_pred_21 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

#choose k = 27
wbcd_test_pred_27 = knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)



#compare
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)



```







